### 2016-09-16 manuscript/manuscript.tex
Update to ITHIM manuscript #1

\documentclass{bioinfo}
\usepackage{url}
\copyrightyear{2016}
\pubyear{2016}

\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\package}{\emph{ITHIM}}
\newcommand{\ithim}{\emph{ITHIM}}
\newcommand{\R}{\emph{R}}
\newcommand{\sgy}[1]{{\itshape\col{#1}}}
\newcommand{\XX}{\sgy{XX}}
\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
\newcommand{\af}{\rho}
\newcommand{\mets}{MET-hrs./week}
\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
\newcommand{\vmt}{$\mathrm{VMT}$}

\begin{document}
\firstpage{1}

\title[\package{}]{\package{}: an \R{} package for implemetation of
  the ITHIM model}
\author[Younkin \textit{et~al}]{Samuel G. Younkin$^{1}$,
  Jason Vargo$^{1}$\footnote{to whom correspondence should be addressed},
  $\ldots$
  \ and Jonathan Patz$^{1}$}
\address{$^{1}$Global Health Institute\\
University of Wisconsin{\textendash}Madison, Madison, WI USA\\
}
\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle

\begin{abstract}

\section{Summary:}
\col[blue]{While increased active transportation, primarily cycling
  and walk, is often held up as a goal for improved public health —
  via reduced chronic disease, improved air quality, and reductions in
  greenhouse gas emissions --- obtaining estimates for the quantiative
  impacts of such shifts remains infrequent, in not altogether
  absent. Using U.S. metropolitan counties from the 10 regions defined
  by the Department of Health and Human Services (HHS1-10) we perform
  a comparative risk analysis between each regions and national
  averages as the baseline.  We investigated the reduced chronic
  disease burden of (x,y,z outcomes) from increased physical activity
  from population shifts in the amount of biking and walking doen for
  transport.  Our methods employ the Integrated Transport and Health
  Impacts Model (ITHIM) informed by regional travel parameters using
  the National Household Transportation Survey, non-travel activity
  metircs from the American Time Use Survey, and baseline Disease
  burden from the Centers for Disease Control and Prevention's WONDER
  database.  Our findings show that the West coast and the Northeast
  of the U.S. have travel patterns that, if mirrored by the nation,
  would save the U.S. thousands to hundreds of thousands DALYs.}

\section{Availability:}
\col[blue]{\package{} may be explored using the user interface
  avaiable at \webpage{}.  The source code and data files for
  \package{} are publicly available on GitHub at \webpage{}.}

\section{Supplementary information:}

\col[blue]{See \webpage{} for links to the source code, manual,
  tutorial and links for more information about the ITHIM model.}

\section{Contact:}
\col[blue]{\href{javargo@wisc.edu}{javargo@wisc.edu}}
\end{abstract}

\section{Introduction}

\col{ITHIM is a statistical model that integrates data on travel
  patterns, physical activity, fine particulate matter, GHG emissions,
  and disease and injuries. Based on population and travel
  scenarios. The model has been used to calculate the health impacts
  of walking and bicycling short distances usually traveled by car or
  driving low-emission automobiles \cite{woodcock2013,maizlish2013}.}

\col{The ITHIM model uses comparative risk assessment through which it
  formulates a change in the disease burden, resulting from the shift
  in the exposure distribution from a baseline scenario to an
  alternative scenario.}

\col{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
  eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim
  ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut
  aliquip ex ea commodo consequat. Duis aute irure dolor in
  reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
  pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
  culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
  dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
  incididunt ut labore et dolore magna aliqua.}

\col{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
  eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim
  ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut
  aliquip ex ea commodo consequat. Duis aute irure dolor in
  reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
  pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
  culpa qui officia deserunt mollit anim id est laborum.}

\begin{methods}

\section{Methods}

The ITHIM model uses the comparative risk assesment framework to
estimate the population attributable fraction, $\af$.  In our case
$\af$ represents the proportional change in mortality and morbidity if
the physical activity levels were adjusted to represent an alternate
scenario.

\begin{equation}
\af = \frac{\int R(x)P(x) dx  - \int R(x)Q(x) dx }{\int R(x)P(x) dx}
\end{equation}

Here $R(x)$ represents the risk when the individual has exposure $x$,
in our case physical activity (\mets).  $P(x)$ and $Q(x)$ denote the
population density of individuals with exposure $x$, in the baseline
and alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of travel-related exposure and non-travel-related
exposure.

\subsection{Active Transport}

The time spent walking or cycling, i.e., active transport time, is
modeled as a log-normal random variable with means defined for each
age-sex stratum.  The non-travel-related physical activity is also
modeled as log-normal with stratum-specific means.  Estimates for
non-travel-related physical activity are diffiuclt.  We used the
American Time Use Survey data to estimate the relative means for each
of the age-sex strata.  We then treat the mean of the referent group
as a variable and compute results over a grid of values for this
variable.  See Figure \ref{dalyFigure}.

The travel-related exposure is estimated using active transport time
along with assumptions about how much physical activity is required
for cycling and walking.  Active transport time and non-travel-related
physical activity are modeled as independent random variables.  A
fixed value was used as the coefficient of variation for the
distributin of active transport time. A different, but also fixed,
value was used for the non-travel-related physical activity
distribution.

\subsection{Road Injuries}

\col[red]{\textit{
We chose to model road injuries in a more simplifed manner than the
authors of ITHIM.  The original ITHIM requires information about the
transportational infrastructure that cannot be accurately modeled on a
regional level, and is more well-suited to say a collection of
counties as a case study. \cite{maizlish2013}  We use the overall percent
of travel by walking and road fatalities per U.S. metropolitan area to
estimate road fatalaties in U.S. metropolitan areas given the percent
of travel by walking. \cite{UWTechReport}
}}

\col[red]{\textit{
We need to estimate the increase/decrease in road injuries/fatalaties
nation-wide as walking/cycling time increases.  We need a relative
risk of road fatalaty as a function of active transport time...
}}

\col[red]{\textit{

}}

\subsection{Air Pollution}

Can we model the decrease in \pEm{2.5} nationally as a function of
decrease in driving time?  The original plan was to use a baseline
value for \pEm{2.5} and a linear relationship bewteen change in
\pEm{2.5} and change in \vmt{}.

\end{methods}

\section{U.S. Metro-Regional Analysis}
\col[blue]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
occaecat cupidatat non proident, sunt in culpa qui officia deserunt
mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
ullamco laboris nisi ut aliquip ex ea commodo consequat.}

\subsection{Data Sources}

\col[red]{Lorem ipsum dolor sit amet, consectetur adipisicing)) elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
occaecat cupidatat non proident, sunt in culpa qui officia deserunt
mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
ullamco laboris nisi ut aliquip ex ea commodo consequat.}

\begin{figure}[t]
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/HHSmap}}
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
    \caption{Change in total national DALYs for each regional
      transportation estimates.}\label{dalyFigure}
\end{figure}

\section*{Acknowledgment}
\paragraph{Funding\textcolon} SGY is supported by$\ldots$ , JV, JP, $\ldots$

%\bibliographystyle{natbib}
\bibliographystyle{bmc_article}
%\bibliographystyle{mypapers}
\bibliography{ITHIM}
\end{document}