### 2016-10-05 manuscript/manuscript.tex

\documentclass{bioinfo}
\usepackage{url}
\copyrightyear{2016}
\pubyear{2016}

\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\package}{\emph{ITHIM}}
\newcommand{\ithim}{\emph{ITHIM}}
\newcommand{\R}{\emph{R}}
\newcommand{\sgy}[1]{{\itshape\col{#1}}}
\newcommand{\XX}{\sgy{XX}}
\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
\newcommand{\af}{\rho}
\newcommand{\mets}{MET-hrs./week}
\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
\newcommand{\vmt}{$\mathrm{VMT}$}

\begin{document}
\firstpage{1}

\title[\package{}]{A comparative risk assesment of U.S. regional
  active transport using the ITHIM model}
\author[Younkin \textit{et~al}]{Samuel G. Younkin$^{1}$,
  Jason Vargo$^{1}$\footnote{to whom correspondence should be addressed},
  $\ldots$
  \ and Jonathan Patz$^{1}$}
\address{$^{1}$Global Health Institute\\
University of Wisconsin{\textendash}Madison, Madison, WI USA\\
}
\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle

\begin{abstract}

\section{Summary:}
\col[blue]{While increased active transportation, primarily cycling
  and walking, is often held up as a goal for improved public health —
  via reduced chronic disease, improved air quality, and reductions in
  greenhouse gas emissions — obtaining estimates for the quantiative impacts of such shifts remains infrequent, if not altogether absent from related policy action. Using U.S. metropolitan counties from seven states we perform a comparative risk analysis between each regions and national averages as the baseline.  We employ a portion of the Integrated Transport and Health Impacts Model (ITHIM) to investigate the reduced chronic disease burden of (x,y,z outcomes) from changing levels of physical activity with population shifts in the amount of biking and walking done for transport. The analysis is informed by regional travel parameters from the National Household Transportation Survey, non-travel activity metrics from the American Time Use Survey, and baseline disease burden from the Centers for Disease Control and Prevention's WONDER database and the Global Burden of Disease dataset.  Our findings show that the California and New York State have travel patterns that, if mirrored by the nation, would save the U.S. upwards of fifty thousand disability adjusted life years.}

\section{Availability:}
\col[blue]{\package{} may be explored using the user interface
  avaiable at \webpage{}.  The source code and data files for
  \package{} are publicly available on GitHub at \webpage{}.}

\section{Supplementary information:}

\col[blue]{See \webpage{} for links to the source code, manual,
  tutorial and links for more information about the ITHIM model.}

\section{Contact:}
\col[blue]{\href{javargo@wisc.edu}{javargo@wisc.edu}}
\end{abstract}

\section{Introduction}

\col{Regular exercise has been shown to provide health benefits in terms of chronic disease prevention \cite{warburton2006}, cognitive function \cite{hillman2008}, and overall well being and satisfaction with life \cite{maher2013}. Still only half of Americans obtain the recommended levels of aerobic activity, which is 2.5 hours per week. There is considerable variability in the attainment of this goal by location within the country. Supportive local built and policy environments are important determinants of population-level activity patterns. Particularly, as they facilitate and foster routine moderate physical activity, such as walking, local decissions on design and policy can play a huge role in the Nation's health.}

\col{To better understand the potential for attainable shifts in transportation behavior across the US, we employed a comparative risk assessment methodology to compare national averages for time spent walking and cycling for transport to those of specific states. Our estimates can provide an important evidence base to encourage new construction, retrofits, land use policy, and programatic funding to support active transportation. Particularly in places with poor attainmnet of the physical activity guidelines, such investments can be importnat pieces of lareger population health interventions that have effective and long-lasting impacts. }

\col{ITHIM is a statistical model that integrates data on travel
  patterns, physical activity, fine particulate matter, GHG emissions,
  and disease and injuries. Based on population and travel
  scenarios. The model has been used to calculate the health impacts
  of walking and bicycling short distances usually traveled by car or
  driving low-emission automobiles \cite{woodcock2013,maizlish2013}.}

\col{The ITHIM model uses comparative risk assessment through which it
  formulates a change in the disease burden, resulting from the shift
  in the exposure distribution from a baseline scenario to an
  alternative scenario.}

\begin{methods}

\section{Methods}

\col[blue]{\textit{
    The ITHIM model uses the comparative risk
    assesment framework to estimate the population attributable
    fraction, $\af$.  In our case $\af$ represents the proportional
    change in mortality and morbidity if the physical activity levels
    were adjusted to represent an alternate scenario.
}}

\begin{equation}
\af = \frac{\int R(x)P(x) dx  - \int R(x)Q(x) dx }{\int R(x)P(x) dx}
\end{equation}

\col[blue]{\textit{
    Here $R(x)$ represents the risk when the individual has exposure $x$,
in our case physical activity (\mets).  $P(x)$ and $Q(x)$ denote the
population density of individuals with exposure $x$, in the baseline
and alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of travel-related exposure and non-travel-related
exposure.
}}

\subsection{Active Transport}

\col[blue]{\textit{The time spent walking or cycling, i.e., active
    transport time, is modeled as a log-normal random variable with
    means defined for each age-sex stratum.  The non-travel-related
    physical activity is also modeled as log-normal with
    stratum-specific means.}}

\col[blue]{\textit{
The travel-related exposure is estimated using active transport time
along with assumptions about how much physical activity is required
for cycling and walking.  Active transport time and non-travel-related
physical activity are modeled as independent random variables.  A
fixed value was used as the coefficient of variation for the
distributin of active transport time. A different, but also fixed,
value was used for the non-travel-related physical activity
distribution.
}}

\subsection{Road Injuries}

\col[red]{\textit{
We chose to model road injuries in a more simplifed manner than the
authors of ITHIM.  The original ITHIM requires information about the
transportational infrastructure that cannot be accurately modeled on a
regional level, and is more well-suited to say a collection of
counties as a case study. \cite{maizlish2013}  We use the overall percent
of travel by walking and road fatalities per U.S. metropolitan area to
estimate road fatalaties in U.S. metropolitan areas given the percent
of travel by walking. \cite{UWTechReport}
}}

\col[red]{\textit{
We need to estimate the increase/decrease in road injuries/fatalaties
nation-wide as walking/cycling time increases.  We need a relative
risk of road fatalaty as a function of active transport time...
}}

\subsection{Air Pollution}

\col[red]{\textit{
    Can we model the decrease in \pEm{2.5} nationally as a function of
decrease in driving time?  The original plan was to use a baseline
value for \pEm{2.5} and a linear relationship bewteen change in
\pEm{2.5} and change in \vmt{}.
}}

\end{methods}

\begin{figure}[t]
    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig1}}
    \caption{Map of HHS Regions and metropolitan areas}\label{HHSmap}
\end{figure}

\section{U.S. Metro-Regional Analysis}

\begin{figure}[t]
    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
    \caption{Change in total national DALYs for each regional
      transportation estimates.}\label{dalyFigure}
\end{figure}

\col[blue]{\begin{itemize}
\item We used the National Household Transportation Survey \cite{NHTS} to
  estimate the mean walk and cycle times for individuals that live in
  metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
  corresponds to individuals whose homes are in an MSA or cMSA)
  \bi\item Data quality issues along with assumptions about travel speed
created mirky data.  The median values of trip time were used to avoid
undue influence by outliers.\ei
\item We then divided these into the 10 regions of the U.S. defined by
  the Department of Health and Human Services (HH1-10)
\item Using only the active transport component of the ITHIM model we
  ran comparative analyses for each of the ten regions, using the
  national averages as the baseline.
\item We used data from the American Time Use Survey \cite{ATUS} to
  estimate the relative means for each of the age-sex strata.  We then
  treat the mean of the referent group as a variable and compute
  results over a grid of values for this variable.  See Figure
  \ref{dalyFigure}.
\item Values for DALYs were found using the CDC Wonder database \cite{CDCWonder}
\end{itemize}}

\begin{figure}[t]
  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
    \caption{Mean travel time by mode (walking/cycling) for each of
      the four regions with a decrease in DALYs
      $(\textrm{HHS01},\textrm{HHS02},\textrm{HHS09},\textrm{HHS10})$. The
      national means are displayed on the far right.  These four
      regions correspond to New England/NewYork and the West
      Coast.}\label{meanMatrices}
\end{figure}

\section*{Acknowledgment}
\paragraph{Funding\textcolon} \col[red]{SGY is supported by$\ldots$ , JV, JP, $\ldots$}

%\bibliographystyle{natbib}
\bibliographystyle{bmc_article}
%\bibliographystyle{mypapers}
\bibliography{ITHIM}
\end{document}