### 2016-10-05 manuscript/manuscript.tex
Edits to manuscript.

\documentclass{bioinfo}
\usepackage{url}
\copyrightyear{2016}
\pubyear{2016}

\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\package}{\emph{ITHIM}}
\newcommand{\ithim}{\emph{ITHIM}}
\newcommand{\R}{\emph{R}}
\newcommand{\sgy}[1]{{\itshape\col{#1}}}
\newcommand{\XX}{\sgy{XX}}
\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
\newcommand{\af}{\textrm{PAF}}
\newcommand{\mets}{MET-hrs./week}
\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
\newcommand{\vmt}{$\mathrm{VMT}$}
\newcommand{\logNormal}{log-normal}

\begin{document}
\firstpage{1}

\title[\package{}]{A comparative risk assesment of U.S. regional
  active transport using the ITHIM model}
\author[Younkin \textit{et~al}]{Samuel G. Younkin$^{1}$,
  Jason Vargo$^{1}$\footnote{to whom correspondence should be addressed},
  $\ldots$
  \ and Jonathan Patz$^{1}$}
\address{$^{1}$Global Health Institute\\
University of Wisconsin{\textendash}Madison, Madison, WI USA\\
}
\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle

\begin{abstract}

\section{Summary:}

While increased active transportation, primarily cycling and walking,
is often held up as a goal for improved public health --- via reduced
chronic disease, improved air quality, and reductions in greenhouse
gas emissions — obtaining estimates for the quantiative impacts of
such shifts remains infrequent, if not altogether absent from related
policy action. Using U.S. metropolitan counties from seven states we
perform a comparative risk analysis between each regions and national
averages as the baseline.  We employ a portion of the Integrated
Transport and Health Impacts Model (ITHIM) to investigate the reduced
chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
levels of physical activity with population shifts in the amount of
biking and walking done for transport. The analysis is informed by
regional travel parameters from the National Household Transportation
Survey, non-travel activity metrics from the American Time Use Survey,
and baseline disease burden from the Centers for Disease Control and
Prevention's WONDER database and the Global Burden of Disease dataset.
Our findings show that the California and New York State have travel
patterns that, if mirrored by the nation, would save the U.S. upwards
of fifty thousand disability adjusted life years.

\section{Availability:}

\package{} may be explored using the user interface avaiable
at \webpage{}.  The source code and data files for \package{} are
publicly available on GitHub at \webpage{}.

\section{Supplementary information:}

See \webpage{} for links to the source code, manual, tutorial and
links for more information about the ITHIM model.

\section{Contact:}
\href{javargo@wisc.edu}{javargo@wisc.edu}

\end{abstract}

\section{Introduction}

Regular exercise has been shown to provide health benefits in terms of
chronic disease prevention \cite{warburton2006}, cognitive function
\cite{hillman2008}, and overall well being and satisfaction with life
\cite{maher2013}. Still only half of Americans obtain the recommended
levels of aerobic activity, which is 2.5 hours per week. There is
considerable variability in the attainment of this goal by location
within the country. Supportive local built and policy environments are
important determinants of population-level activity
patterns. Particularly, as they facilitate and foster routine moderate
physical activity, such as walking, local decissions on design and
policy can play a huge role in the Nation's health.

To better understand the potential for attainable shifts in
transportation behavior across the US, we employed a comparative risk
assessment methodology to compare national averages for time spent
walking and cycling for transport to those of specific states. Our
estimates can provide an important evidence base to encourage new
construction, retrofits, land use policy, and programatic funding to
support active transportation. Particularly in places with poor
attainmnet of the physical activity guidelines, such investments can
be importnat pieces of lareger population health interventions that
have effective and long-lasting impacts.

ITHIM is a statistical model that integrates data on travel patterns,
physical activity, fine particulate matter, GHG emissions, and disease
and injuries. Based on population and travel scenarios. The model has
been used to calculate the health impacts of walking and bicycling
short distances usually traveled by car or driving low-emission
automobiles \cite{woodcock2013,maizlish2013}.

The ITHIM model uses comparative risk assessment through which it
formulates a change in the disease burden, resulting from the shift in
the exposure distribution from a baseline scenario to an alternative
scenario.

\begin{methods}

\section{Methods}

The ITHIM model uses the comparative risk assesment framework to
estimate the population attributable fraction, $\af$.  In our case the
$\af$ represents the proportional change in national disease burden if
age and sex-specific active transport time means were changed.

\begin{equation}
\af = \frac{\int R(x)P(x) dx  - \int R(x)Q(x) dx }{\int R(x)P(x) dx}
\end{equation}

Here $R(x)$ represents the risk when the individual has exposure $x$,
physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
density of individuals with exposure $x$, in the baseline and
alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of two independent random variables, travel-related
exposure and non-travel-related exposure.  Non-travel-related exposure
is modeled with a \logNormal{} distribution with the age-sex specific
mean ratios for non-travel-related exposure are estimated from the
ATUS \cite{ATUS}.  The referent group mean is treated as a variable
and the coefficient of variation is treated as constant across age-sex
strata and estimated with ATUS.  We see in Figure \ref{dalyFigure}
that varying the mean values for non-travel-related exposure does not
greatly affect the estimate for disease burden.  The distribution for
travel-related exposure is estimated from the time spent walking or
cycling, i.e., active transport time, which as in the original ITHIM
model has a \logNormal{} distribution with constant coefficient of
variation across age-sex strata \cite{}.  Travel-related exposure
is estimated from active transport time using assumptions about how
much physical activity is required for cycling and walking.

\end{methods}

\begin{figure}[t]
  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
    \caption{California \& New York Averted DALYs Nationwide}\label{California}
\end{figure}

\section{Results}

We used the National Household Transportation Survey \cite{NHTS} to
estimate the mean walk and cycle times for individuals that live in
metropolitan areas for each of seven states (CA, TX, NY, FL, VA, NC
and AZ) which had sufficient sample size for reasonable estimates of
age-sex specific travel-related walking and cycling means, Figure
\ref{meanMatrices}.  Of these seven states two were found to have
active transport behavior that would avert thousands of DALYs if
adopted nation-wide. Values for national disease-specific DALYs were
found using the CDC Wonder database \cite{CDCWonder}.

%% \begin{itemize}
%% \item We used the National Household Transportation Survey \cite{NHTS} to
%%   estimate the mean walk and cycle times for individuals that live in
%%   metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
%%   corresponds to individuals whose homes are in an MSA or cMSA)
%%   \bi\item Data quality issues along with assumptions about travel speed
%% created mirky data.  The median values of trip time were used to avoid
%% undue influence by outliers.\ei
%% \item We then divided these into the 10 regions of the U.S. defined by
%%   the Department of Health and Human Services (HH1-10)
%% \item Using only the active transport component of the ITHIM model we
%%   ran comparative analyses for each of the ten regions, using the
%%   national averages as the baseline.
%% \item We used data from the American Time Use Survey \cite{ATUS} to
%%   estimate the relative means for each of the age-sex strata.  We then
%%   treat the mean of the referent group as a variable and compute
%%   results over a grid of values for this variable.  See Figure
%%   \ref{dalyFigure}.
%% \item 
%% \end{itemize}

\begin{figure}[t]
    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
    \caption{Change in total national DALYs for each regional
      transportation estimates.}\label{dalyFigure}
\end{figure}

\begin{figure}[t]
  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
    \caption{Mean travel time by mode (walking/cycling) for each of
      seven states National means are displayed on the right.}\label{meanMatrices}
\end{figure}

\section*{Acknowledgment}
\paragraph{Funding\textcolon} SGY is supported by \ldots,\ JV, JP by \ldots

%\bibliographystyle{natbib}
\bibliographystyle{bmc_article}
%\bibliographystyle{mypapers}
\bibliography{ITHIM}
\end{document}

%% \subsection{Road Injuries}

%% \textit{ We chose to model road injuries in a more simplifed manner
%%   than the authors of ITHIM.  The original ITHIM requires information
%%   about the transportational infrastructure that cannot be accurately
%%   modeled on a regional level, and is more well-suited to say a
%%   collection of counties as a case study. \cite{maizlish2013} We use
%%   the overall percent of travel by walking and road fatalities per
%%   U.S. metropolitan area to estimate road fatalaties in
%%   U.S. metropolitan areas given the percent of travel by
%%   walking. \cite{UWTechReport}}

%% \textit{ We need to estimate the increase/decrease in road
%%   injuries/fatalaties nation-wide as walking/cycling time increases.
%%   We need a relative risk of road fatalaty as a function of active
%%   transport time\ldots}

%% \subsection{Air Pollution}

%% \textit{ Can we model the decrease in \pEm{2.5} nationally as a
%%   function of decrease in driving time?  The original plan was to use
%%   a baseline value for \pEm{2.5} and a linear relationship bewteen
%%   change in \pEm{2.5} and change in \vmt{}.}
%% \begin{figure}[t]
%%     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig5.pdf}}
%%     \caption{Counts}\label{counts}
%% \end{figure}