### 2016-10-11 methodsPaper/tex/abstract.tex

%% While increased active transportation, primarily cycling and walking,
%% is often held up as a goal for improved public health --- via reduced
%% chronic disease, improved air quality, and reductions in greenhouse
%% gas emissions — obtaining estimates for the quantiative impacts of
%% such shifts remains infrequent, if not altogether absent from related
%% policy action. Using U.S. metropolitan counties from seven states we
%% perform a comparative risk analysis between each regions and national
%% averages as the baseline.  We employ a portion of the Integrated
%% Transport and Health Impacts Model (ITHIM) to investigate the reduced
%% chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
%% levels of physical activity with population shifts in the amount of
%% biking and walking done for transport. The analysis is informed by
%% regional travel parameters from the National Household Transportation
%% Survey, non-travel activity metrics from the American Time Use Survey,
%% and baseline disease burden from the Centers for Disease Control and
%% Prevention's WONDER database and the Global Burden of Disease dataset.
%% Our findings show that the California and New York State have travel
%% patterns that, if mirrored by the nation, would save the U.S. upwards
%% of fifty thousand disability adjusted life years.

%\subsection{Summary:}
%% \section{Availability:}

%% %% \package{} may be explored using the user interface avaiable
%% %% at \webpage{}.
%% The source code and data files for \package{} are publicly available
%% on GitHub via a link on \webpage{}.

% \section{Supplementary information:}

% See \webpage{} for links to the source code, manual, tutorial and
% links for more information about the ITHIM model.

% \section{Contact:}
% \href{javargo@wisc.edu}{javargo@wisc.edu}