### 2016-10-12 methodsPaper/tex/introduction.tex
Tested pucher2010 reference in methods paper
Seems like it worked.

Regular exercise has been shown to provide health benefits in terms of
chronic disease prevention \cite{warburton2006}, cognitive function
\cite{hillman2008}, and overall well being and satisfaction with life
\cite{maher2013}. Still only half of Americans obtain the recommended
levels of aerobic activity, which is 2.5 hours per week. There is
considerable variability in the attainment of this goal by location
within the country. Supportive local built and policy environments are
important determinants of population-level activity
patterns. Particularly, as they facilitate and foster routine moderate
physical activity, such as walking, local decissions on design and
policy can play a huge role in the Nation's health.

%% To better understand the potential for attainable shifts in
%% transportation behavior across the US, we employed a comparative risk
%% assessment methodology to compare national averages for time spent
%% walking and cycling for transport to those of specific states. Our
%% estimates can provide an important evidence base to encourage new
%% construction, retrofits, land use policy, and programatic funding to
%% support active transportation. Particularly in places with poor
%% attainmnet of the physical activity guidelines, such investments can
%% be importnat pieces of lareger population health interventions that
%% have effective and long-lasting impacts.

ITHIM is a statistical model that integrates data on active transport,
physical activity, fine particulate matter and greenhouse gas
emissions to provide estimates for the proportional change in
mortality and morbidity for given baseline and alternate travel
scenarios. The model has been used to calculate the health impacts of
walking and bicycling short distances usually traveled by car or
driving low-emission automobiles \cite{woodcock2013,maizlish2013}.

ITHIM uses a comparative risk assessment framework for the active
transport component of the model.  We improve on the existing
implementation of the active transport component by including the
distribution of non-travel-related activity, improving numerical
precision when computing the population attributable fraction and
creating a simple user-interface for the custom-built \R{}
package \package{}.

Dummy ref \cite{pucher2010}.