### 2016-10-19 methodsPaper/tex/results.tex

Our subset of the US population to only metropolitan counties
consisted of 1,090 of 3,141 (34.7\%) counties, or ???X??? of ???Y???
(???Z???\%) of the total US population. The age and gender
distribution of our subset population was slightly younger (???making
this up???) than the national population. Active transport times
(walking and cycling) for baseline, low, and high scenarios are shown
in \ref{fig2}. Overall mean walking time for high and low
scenarios were ???X??? (min/week) higher and ???Y??? (min/week) lower
than baseline, respectively. Baseline non-travel activity time was
estimated to be ???X??? (min/week) from the ATUS for US metropolitan
populations.

Comparing baseline to low/high scenarios produced estimates of ???X???
change in the disease burden (1,000s DALYs) for our US population
subset. We then varied the mean non-travel activity time and active
travel time to observe how this chaged results (\ref{fig1}). We
observe the most dramatic responses in disease burden estimates to
changes in active travel time when non-travel activity times are
lowest. This result is consistent with the non-linear dose-response
curve for physical activity and several health outcomes. The total
health benefit of increased active travel time (high scenario) varies
from ???X??? to ???Y??? with changing mean non-travel activity time (0
to 25 minutes).

Adapting ITHIM's methods to the R platform enables the rapid and
thorough assessment of parameter influence within the model. Another
important improvement enabled by the shift to a more computationaly
robust platform was to examine methodological assumptions of former
implementations. One critical assumption invovles the approximation of
??? used to calculate the attirbutable fraction (see
Figure \ref{fig:overview}). Previous implementations used quitiles to
approximate this value, essentailly the area under the ???? curve. We
improved upon those methods to use percentiles or 100 points to
estimate the same values. We find our estimates to be much more stable
and reliable at all values of active travel and non-travel activity
parameters.

As exhibited in Figure \ref{fig1}, the variability in disease burden
benfits varies to a much greater degree when using fewer points in the
attirbutable fraction approximation.