### 2016-10-31 methodsPaper/tex/abstract.tex
Edits to Abstract

Comparative risk assessment analyses to quantify the health potential
of active transportation, e.g., walking and cycling, have been
performed in the United States on a limited basis. Current
methodologies, which rely on extensive data collection and user
expertise, have estimated health benefits for select metropolitan
areas under a few alternative scenarios. These analyses have largely
been conducted using the Integrated Transport Health Impacts Model
(ITHIM) which was developed in the UK and has been adapted for use in
the US, primarily California. Here we present an implementation of the
active transportation component of ITHIM to model the health impacts
of modified active transportation levels. Using the newly developed
software package \package{} written in R, along with publicly
available national datasets, we quantify the change in disability
adjusted life years over a wide range of scenarios.  The R package
implemention of ITHIM's physical activity component provides a robust
environment for exploring the relationship between active
transportation and positive health outcomes.  A web-based
user-interface is available for users who want a first estimate for
the impact active transportation has on their
state/county/tract/region.

%% While increased active transportation, primarily cycling and walking,
%% is often held up as a goal for improved public health --- via reduced
%% chronic disease, improved air quality, and reductions in greenhouse
%% gas emissions — obtaining estimates for the quantiative impacts of
%% such shifts remains infrequent, if not altogether absent from related
%% policy action. Using U.S. metropolitan counties from seven states we
%% perform a comparative risk analysis between each regions and national
%% averages as the baseline.  We employ a portion of the Integrated
%% Transport and Health Impacts Model (ITHIM) to investigate the reduced
%% chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
%% levels of physical activity with population shifts in the amount of
%% biking and walking done for transport. The analysis is informed by
%% regional travel parameters from the National Household Transportation
%% Survey, non-travel activity metrics from the American Time Use Survey,
%% and baseline disease burden from the Centers for Disease Control and
%% Prevention's WONDER database and the Global Burden of Disease dataset.
%% Our findings show that the California and New York State have travel
%% patterns that, if mirrored by the nation, would save the U.S. upwards
%% of fifty thousand disability adjusted life years.

%\subsection{Summary:}
%% \section{Availability:}

%% %% \package{} may be explored using the user interface avaiable
%% %% at \webpage{}.
%% The source code and data files for \package{} are publicly available
%% on GitHub via a link on \webpage{}.

% \section{Supplementary information:}

% See \webpage{} for links to the source code, manual, tutorial and
% links for more information about the ITHIM model.

% \section{Contact:}
% \href{javargo@wisc.edu}{javargo@wisc.edu}