### 2016-11-23 R/ITHIM.Rmd
Now methods paper is created with knitr
See files ./lib/*.Rnw for editing content.

# Integrated Transport and Health Impacts Model

## R Package _ITHIM_
<!-- This document will eventually serve as a compendium to the R package
_ITHIM_.  It will describe how to use the package to implement the
ITHIM model, as well as a detailed description of the model itself.

As a working document it will serve as a place for SGY to post
results, questions, interpreations, notes, etc.-->

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
#install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
library("ggplot2")
library("devtools")
library("reshape2")
library("plyr")
library("dplyr")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
document("~/ITHIM/")
install("~/ITHIM/")
#install_github("syounkin/ITHIM", ref="devel")
install_github("smbache/magrittr", ref="master")

library("magrittr")
library("ITHIM")

fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

## HHS Metro Region Parameter Sets

First we use NHTS to estimate mean walking and cycling times.

### NHTS

```{r compute, echo = FALSE, eval = TRUE}
compute <- FALSE
```

#### Create a handful of functions to help parse data

```{r parsely, eval = TRUE, echo = FALSE}
convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

countPeople <- function(data){
 nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
}

tripTime <- function(data, mode = "walk"){
 return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
}

reformatData <- function(data){
foo <- matrix(unlist(data), nrow = 7, ncol = 2)
foo <- rbind( c(0, 0), foo)
dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
return(foo[,c("M","F")])
}

```

#### Read Two of the Four NHTS Data Tables

```{r NHTS, eval = compute, echo = FALSE}
DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")

NHTS.trips <- with(DAYV2PUB,{
    data.frame(HOUSEID = HOUSEID,
                   PERSONID = PERSONID,
                   TDCASEID = TDCASEID,
                   AT = ifelse(TRPTRANS == "22", "cycle",
                        ifelse(TRPTRANS == "23", "walk",
                        ifelse(!is.na(TRPTRANS), "other", NA))),
                   TRVL_MIN = as.numeric(TRVL_MIN),
                   #TRVLCMIN = as.numeric(TRVLCMIN),
                   URBRUR  = as.factor(URBRUR),
                   MSACAT = as.factor(MSACAT),
                   MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                   TRPTRANS = as.factor(TRPTRANS),
                   TRPMILES = as.numeric(TRPMILES)
)})

PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9")
names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")

NHTS.person <- with(PERV2PUB,{
                    data.frame(HOUSEID = HOUSEID,
                               PERSONID = PERSONID,
                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
                               AGE = as.factor(convertToAgeClass(R_AGE)),
                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                               HHSTATE = as.character(HHSTATE),
                               HHS = as.character(convertToHHS(as.character(HHSTATE))), stringsAsFactors = FALSE

)
})

```

#### Subset and Merge

We now correct for zero trip times in NHTS.

```{r subset, eval = compute, fig.width = 10, echo = FALSE}
NHTS.person <- subset( NHTS.person, MSA == TRUE )
NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))

NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
saveRDS(NHTS, file = "./data/NHTS.rds")
```

```{r loadNHTS, eval = !compute}
NHTS <- readRDS(file = "./data/NHTS.rds")
```

#### Compute National Metro Average

```{r natl, eval = TRUE, echo = FALSE, warning = FALSE}
listOfLists <- lapply(split(NHTS, NHTS$SEX),function(x) split(x, x$AGE))
counts <- lapply(listOfLists, function(x) lapply(x, countPeople))
walkLength <- lapply(listOfLists, function(x) lapply( x, tripTime, mode = "walk"))
cycleLength <- lapply(listOfLists, function(x) lapply( x, tripTime, mode = "cycle"))

countMatrix <- reformatData(counts)
walkMatrix <- reformatData(walkLength)
cycleMatrix <- reformatData(cycleLength)

meanWalkList <- walkMatrix/countMatrix*7
meanCycleList <- cycleMatrix/countMatrix*7

RwtList <- ifelse(meanWalkList == 0, 1e-4, meanWalkList)
RctList <- ifelse(meanCycleList == 0, 1e-4, meanCycleList)

RwtNational <- RwtList/RwtList[3,2]
RctNational <- RctList/RctList[3,2]
muwtNational <- meanWalkList[3,2]
muctNational <- meanCycleList[3,2]

rm("listOfLists","counts","walkLength","cycleLength","countMatrix","walkMatrix","cycleMatrix","meanWalkList","meanCycleList","RwtList","RctList","muwtList","muctList")
```

#### Compute Mean Trip Lengths

```{r count, eval = TRUE, echo = FALSE}
listOfLists <- lapply(split(NHTS, NHTS$HHSTATE),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
countMatrix <- lapply(counts, reformatData)
walkMatrix <- lapply(walkLength, reformatData)
cycleMatrix <- lapply(cycleLength, reformatData)
meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)
meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)

zeroReferentCase <- function(x){
  if(x[3,2]==0){
    return(matrix(1e-3,nrow=nrow(x),ncol=ncol(x), dimnames = dimnames(x)))
  }else{
    return(ifelse(x==0, 1e-3, x/x[3,2]))
  }
}

RwtList <- lapply( meanWalkList, zeroReferentCase)
RctList <- lapply( meanCycleList, zeroReferentCase)
muwtList <- lapply( meanWalkList, function(x) ifelse(x[3,2]==0,1e-3, x[3,2]))
muctList <- lapply( meanCycleList, function(x) ifelse(x[3,2]==0,1e-3, x[3,2]))
```

## Non-Travel METs

```{r newMETs, echo = TRUE, eval = TRUE, echo = FALSE}
newMETsFile <- "ATUS_to_METS_key.csv"

newMETsFile <- system.file(newMETsFile, package="ITHIM")

newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
dataColumns <- c(1,5)
D <- newMETs[,dataColumns]
metaData <- newMETs[,-dataColumns]
newMETs <- list(Data = D, metaData = metaData)
activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))

rownames(newMETs$Data) <- activityCode
rownames(newMETs$metaData) <- activityCode

newMETs$metaData <- within(newMETs$metaData, {
  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
})

metKey <- newMETs$Data$AinsworthMETS
names(metKey) <- activityCode

```

```{r readnewMETs, echo = TRUE, eval = TRUE, echo = FALSE}
cpsFile <- "~/GHI/data/ATUS/atuscps_2015.dat"
cps <- read.table(cpsFile, header=T, skip=0, sep="," )
cps <- within(cps, {
    TUCASEID <- as.character(TUCASEID)
    TULINENO <- as.character(TULINENO)
    GEREG <- as.factor(GEREG)
    })
cps <- subset(cps, TULINENO == "1")

summaryFile <- "~/GHI/data/ATUS/atussum_2015.dat"
summary <- read.table(summaryFile, header=T, skip=0, sep=",")
metaDataColumns <- 1:24
summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])

summaryList$metaData <- within(summaryList$metaData,{
    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
    AGECAT <- ifelse(TEAGE <5, "00-04",
                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
                            )))))))})

```

We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.

```{r metKey, echo = TRUE, eval = TRUE, echo = FALSE}
allActivities <- colnames(summaryList$Data)
nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]

activities <- nonTravelActivities

metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
names(metKeyVector) <- activities
metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
dimnames(metKeyMatrix) <- list( NULL, activities)

METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )

D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))

```

```{r parameters, eval = TRUE, echo = FALSE}
parameters <- ddply(D, .(SEX, AGECAT), summarise,
             mean = mean(METs, na.rm=T),
             sd = sd(METs, na.rm=T)
             )

METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))

colnames(METmatrix) <- c(as.character(parameters$SEX[1]),as.character(rev(parameters$SEX)[1]))

METmatrix <- METmatrix[,c("male","female")]

METmatrix <- METmatrix/METmatrix[3,2]
METmatrix[2,] <- METmatrix[3,]
rownames(METmatrix) <- paste0("ageClass",1:8)
colnames(METmatrix) <- c("M","F")
METmatrix
```

## Create List of ITHIM Objects

```{r foobar, eval = TRUE, echo = FALSE}
sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
    return(sum(unlist(data.frame(compareModels(ITHIM.baseline,ITHIM.scenario)$daly.delta)[-1,]))) # AgeClass 1 is removed from totals
    }
```

```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
stateVec <- c("CA","TX","NY","FL") #c("CA","TX","NY","FL","VA","NC","AZ")
stateVecOrdered <- c("TX","FL","NY","CA") #c("NC","TX","VA","FL","AZ","NY","CA")

upper <- 10; length <- 5; wd <- upper/length;

DALYsTotal <- data.frame()
ITHIMList <- list()
ITHIM.default <- createITHIM()
ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravel", 4)
ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravelMatrix", METmatrix)
ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)

ITHIM.national <- ITHIM.default
ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)

for (HHSTATE in stateVec){
    ITHIMList[[HHSTATE]] <- ITHIM.default
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
  }

for( muNonTravel in seq(0, upper, by = wd)){

  rm(DALYs)

  for (HHSTATE in stateVec){
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
  }

  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)

  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
  DALYsTotal <- rbind(DALYsTotal, DALYs)
}

DALYsTotal <- within(DALYsTotal, {
  State <- factor(State, levels = stateVecOrdered)
})

p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))

p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper))

p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()

```

### Visualization of trip time matrices

```{r meanMatrices, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}

ITHIMList <- c(ITHIMList, list("Natl." = ITHIM.national))

meanMatrix <- function(ITHIM, mode = "Walking"){
   if(mode == "Walking"){
      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
   }else if (mode == "Cycling"){
      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
   }else{
      message("error in meanMatrix function")
   }
}

meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "Walking")
meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "Cycling")

D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
names(D) <- c("ageClass", "mode", "Sex", "time", "State")

D <- within(D,{
  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
  Sex <- factor(Sex, levels = c("F","M"))
})

p <- ggplot(subset(D,ageClass != 1), aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ State, scales = "free_y") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 5))

```

```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
ITHIM.national <- updateITHIM(ITHIM.national, parName = "muNonTravel", 10)

ITHIM.CA <- ITHIMList[["CA"]]
ITHIM.CA <- updateITHIM(ITHIM.CA, parName = "muNonTravel", 10)
foo1 <- compareModels(ITHIM.national,ITHIM.CA)
foobar1 <- data.frame(cbind(melt(foo1$daly.delta),paste0("ageClass",1:8)),State = "CA")

foobar <- foobar1
names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")

ITHIM.NY <- ITHIMList[["NY"]]
ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
foo2 <- compareModels(ITHIM.national,ITHIM.NY)
foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
foobar <- rbind(foobar1,foobar2)
names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")

foobar <- within(foobar, {
  Disease[Disease == "BreastCancer"] <- "Breast Cancer"
  Disease[Disease == "ColonCancer"] <- "Colon Cancer"
  Disease <- factor(Disease, levels = c("Breast Cancer", "Colon Cancer", "Dementia", "Depression", "Diabetes", "CVD"))
})

p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = State))

p + geom_bar(stat="identity", position = "dodge") +
theme(axis.title=element_blank()) +
facet_grid( Disease ~ Sex, scales = "free" ) +
scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") +
scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) +
theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6)) +
scale_fill_manual(values=c("#CC6666", "#9999CC"))

```

```{r counts, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
D <- melt(counts)
names(D) <- c("count","AgeClass","Sex", "State")
D <- within(D,{
  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
  Sex <- factor(Sex, levels = c("F","M"))
})
p <- ggplot(subset(D,AgeClass != "00-04"), aes(AgeClass, count/1e3))
p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Individuals Sampled (Thousands)") + facet_grid( State ~ .) + theme_bw() + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6))
```

## Figure 1

```{r methodsPaper, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)

ITHIM.national <- ITHIM.default
ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)

ITHIM.scenario <- ITHIM.national
ITHIM.national.quints <- updateITHIM(ITHIM.national, "quantiles", seq(1,9,by=1)/10)
ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario, "quantiles", seq(1,9,by=1)/10)

pWalk <- ITHIM.national$means$pWalk[3,2]

DALYs <- c(); DALYs.quints <- c();
atVec <- seq(45, 120, length.out = 10)
muNonTravelVec <- c(1, 10, 25)
DTotal <- data.frame()

for(at in atVec ){  # minutes active transport/week
  for( muNonTravel in muNonTravelVec ){

    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muwt", at*pWalk)
    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muct", at*(1-pWalk))
    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muNonTravel", muNonTravel)

    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muwt", at*pWalk)
    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muct", at*(1-pWalk))
    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muNonTravel", muNonTravel)

    ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)

    DALYs <- sumDALY(ITHIM.national, ITHIM.scenario)
    DALYs.quints <- sumDALY(ITHIM.national.quints, ITHIM.scenario.quints)

    D <- data.frame(quants = c("percentiles","quintiles"), DALYs = c(DALYs,DALYs.quints),at = at, muNonTravel = muNonTravel)

    DTotal <- rbind(D,DTotal)

  }
}

DTotal <- within(DTotal, muNonTravel <- factor(muNonTravel, levels = as.character(muNonTravelVec), labels = c("low","medium","high")))
DTotal <- subset(DTotal, quants == "percentiles")

p <- ggplot(DTotal, aes(x = at, y = -DALYs/1e3, colour = muNonTravel))
p + geom_line() + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)")  + scale_colour_discrete(name="Baseline Activity") #+ theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8))

```

## Figure 2

```{r displayMeans, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}

ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", 10)

D <- rbind(getWalkTime(ITHIM.national),getWalkSpeed(ITHIM.national),getCycleTime(ITHIM.national),getNonTravelMETs(ITHIM.national))
D <- within(D, sex <- factor(sex, levels = c("F","M")))

p <- ggplot(D, aes(x = ageClass, y = mu, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( type ~ ., scales = "free")
```

## Figure 3

```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 1.2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
D <- melt(ITHIM.national$parameters$GBD)
names(D) <- c("disease","sex","ageClass","type","burden", "sex2", "disease2")
D <- within(D, sex <- factor(sex, levels = c("F","M")))

p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( . ~ disease, scales = "free_y")
```

```{r doseresponse, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
RR1vec <- unique(unlist(createActiveTransportRRs()))
lapply(createActiveTransportRRs(),function(x) lapply(x, function(x) unique(c(x))))
#~~~~~~~~~~~~~~~~~~~~~~~
# Manual bull$%#@!
#~~~~~~~~~~~~~~~~~~~~~~~
RR1vec <- RR1vec[RR1vec != 1]
names(RR1vec) <- c("BreastCancer","ColonCancer.M","ColonCancer.F", "CVD", "Dementia", "Depression.young", "Depression.old", "Diabetes")

RRx <- function(x, RR1, k) exp(log(RR1)*x^k)
x <- seq(0,25,0.1)
RRmat <- outer(x, RR1vec, RRx, k = 0.5)
rownames(RRmat) <- x
D <- melt(RRmat)
names(D) <- c("x","disease","RR")

p <- ggplot(D, aes(x = x, y = RR, colour = disease))
p + geom_line()  + theme_bw() + scale_y_continuous(name="Relative Risk") + scale_x_continuous(name="Physical Activity (MET-hrs./week)")
```

### Appendix

Do it again, but with quintiles

```{r ITHIMObjsQuintiles, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
DALYsTotal2 <- data.frame()
ITHIMList <- list()

ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", seq(1, 9, by = 2)/10 )

ITHIM.national <- ITHIM.default
ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)

for (HHSTATE in stateVec){
    ITHIMList[[HHSTATE]] <- ITHIM.default
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
  }

for( muNonTravel in seq(0, upper, by = wd)){

  rm(DALYs)

  for (HHSTATE in stateVec){
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
  }

  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)

  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
  DALYsTotal2 <- rbind(DALYsTotal2, DALYs)
}

DALYsTotal2 <- within(DALYsTotal2, {
  State <- factor(State, levels = stateVecOrdered)
})

DALYsTotal <- rbind(data.frame(DALYsTotal,quints="Percentiles"), data.frame(DALYsTotal2,quints="Quintiles"))

p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))

p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( . ~ quints)

p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( . ~ quints)

```