### 2017-01-23 R/tutorial.Rmd

## ITHIM Tutorial

### Installing the package

```{r figwidth, eval = TRUE,  warning = FALSE, message = FALSE, echo = FALSE}
fig.width <- 6
phi <- (1+sqrt(5))/2 # the golden ratio ;)
options(repos='http://cran.rstudio.com/')
```

First load the package from Github.  To do this you need the package
_devtools_.

```{r, libs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
install.packages("devtools")
library("devtools")
```

Now use _install\_github_ to install _ITHIM_ and be sure to load it with _library_.

```{r libs2, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
library("devtools")
install_github(repo = "syounkin/ITHIM", ref = "master")
library("ITHIM")
```

### Create default ITHIM object
=======
Now that ITHIM is installed and loaded the first step is to create an
ITHIM object that represents the baseline vision.

To do so we may create the default ITHIM object using _createITHIM_.
createITHIM as follows.

```{r create, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
ITHIM.baseline <- createITHIM()
```

### ITHIM Methods

#### getMeans

```{r methods3, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
getMeans(ITHIM.baseline)
```

#### update

```{r update, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
ITHIM.scenario <- update(ITHIM.baseline, list(muwt = 2*getMeans(ITHIM.baseline)$walk))
```

#### compareModels

```{r compareModels, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
CRA.list <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

#### Plot

```{r methods2, eval = FALSE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
plot(ITHIM.baseline)
```

#### getMethodsITHIM

```{r getMethods, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
getMethodsITHIM()
```

### The _ParameterSet_ Object

Each ITHIM object contains a ParameterSet object.

```{r pSet, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
paramSet.baseline <- getParameterSet(ITHIM.baseline)
paramSet.scenario <- getParameterSet(ITHIM.scenario)
=======
An ITHIM object is an R object that conatins, among other things, a
set of parameters that define the model.  To view these parameters we
may use _getParameterSet_.

```{r class, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
getParameterSet(ITHIM.baseline)
```

There are currently fourteen parameters, many of which are matrices,
in the model.  (Recall that we are only using the physical activity
component.)

We have means and mean matrices for walk time, cycle time and
non-travel-related activity and coefficients of variation for active
transport time and non-travel-related activity.  nAgeClass has always
been 8 and meanType is always "referent".  GBD is an array of disease
burden estimates and quantiles is the vector of quantiles used for the
approximation of the PAF integral.

Now to create the scenario we need only change one parameter.
Parameters may be changed with _update_.  First we change the mean
walking time muwt to 120 minutes.

```{r scenario, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
ITHIM.scenario <- update(ITHIM.baseline, list(muwt =120))
>>>>>>> a14bdd7f9847c86f744bf70f3ffe8d2643d2ffb5
```

```{r class, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
paramSet.baseline
```

```{r class2, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
paramSet.scenario
```