### 2017-02-13 R/ICT.Rmd
Cambridge trip files

```{r ICT1, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
library("devtools")
install("~/scenic/")
library("scenic")

prepare_baseline()

```

```{r ICT2, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
# 'inst/extdata/pcycl_baseline2.csv')
pcycl_baseline <- read.csv(system.file("extdata", "pcycl_baseline2.csv", package = "scenic"), header = T)  #cycling probs file
pcycl_baseline <- pcycl_baseline[,-1]
#
# #reads scenarios constants
METcycling <- 6.44
METwalking <- 4.61
METebikes <- 4.50

#Read cycling speeds by age/sex
cyclingspeed <- read.csv(system.file("extdata", "cyclingSpeeds.csv", package = "scenic"),header=T)

# Read Cycling Probabilities into an R Object
probCycling <- read.csv(system.file("extdata", "cycling-probability.csv", package = "scenic"), header = T, as.is = T)

# Convert all probabilities into odss
oddsCycling <- round(probCycling /( 1 - probCycling), 3)

# Convert data.frame into a list
oddsCycling <- unlist(oddsCycling, use.names = F)

#lookup tables for equity=0,1. MUST BE expressed in ODDS!
Pcyc0.eq0 <- oddsCycling[1:4]
Pcyc0.eq1 <- rep(oddsCycling[5], 4)

# Baseline=NTS years 2011-2014 + individuals between 18-84 y.o + not Wales/Scotland
#bl <- readRDS(system.file("extdata", "bl2014_v2.rds", package = "scenic"))

bl <- read.csv(system.file("extdata", "sample_baseline.csv", package = "scenic"), header = T)

# bl = subset(bl, subset = Age_B01ID < 21 & HHoldGOR_B02ID!=10  & HHoldGOR_B02ID!=11)
# bl$Age[bl$Age_B01ID<16] <- '16.59'
# bl$Age[bl$Age_B01ID>=16] <- '60plus'
# bl$Sex[bl$Sex_B01ID==1] <- 'Male'
# bl$Sex[bl$Sex_B01ID==2] <- 'Female'
#
# #bl <- read.csv('bl2012_18_84ag_reduced.csv', header=T, as.is = T)
#
# #IMPORTANT: from database -> ID needs to be deleted, IndividualID renamed to ID.
#
# baseline <- bl
#
# #handle short walks, creating 6x of each
# df <- baseline[baseline$MainMode_B03ID==1,]
# shortwalks <- data.frame()
#
# for (i in 1:6) {
#   shortwalks <- rbind(shortwalks,df)
# }
#
# # Update trip IDs with new IDs
# #shortwalks$TripID <-  c(max(baseline$TripID) + 1:nrow(shortwalks))
#
# baseline <- rbind(baseline,shortwalks)
# baseline <- baseline[order(baseline$ID),]
#
# #add people w/o trips to baseline
# fnotrips  <- readRDS(system.file("extdata", "people_notrips2014_v2.Rds", package = "scenic"))
# #fnotrips  <- read.csv('People_w_NoTrips2012_ENG_v6_anon.csv',header=T,as.is=T)
#
# # Remove 85+ age group + Wales/Scotland
# fnotrips <- subset(fnotrips, subset = Age_B01ID < 21 & HHoldGOR_B02ID!=10 & HHoldGOR_B02ID!=11)
# #fnotrips <- subset(fnotrips, subset = Age_B01ID != 21)
#
# fnotrips$agesex <- ""
#
# fnotrips$Age <- ""
#
# fnotrips$Sex <- ""
#
# fnotrips[(fnotrips$Age_B01ID >= 6 & fnotrips$Age_B01ID <= 15) & fnotrips$Sex_B01ID == 1,]$agesex <- '16.59Male'
#
# fnotrips[(fnotrips$Age_B01ID >= 6 & fnotrips$Age_B01ID <= 15) & fnotrips$Sex_B01ID == 2,]$agesex <- '16.59Female'
#
# fnotrips[fnotrips$Age_B01ID >= 16 & fnotrips$Sex_B01ID == 1,]$agesex <- '60plusMale'
#
# fnotrips[fnotrips$Age_B01ID >= 16 & fnotrips$Sex_B01ID == 2,]$agesex <- '60plusFemale'
#
# fnotrips$Age <- 0
#
# fnotrips[fnotrips$Sex_B01ID == 1,]$Sex <- "Male"
#
# fnotrips[fnotrips$Sex_B01ID == 2,]$Sex <- "Female"
#
#
# fnotrips[(fnotrips$Age_B01ID >= 6 & fnotrips$Age_B01ID <= 15),]$Age <- '16.59'
#
# fnotrips[fnotrips$Age_B01ID >= 16,]$Age <- '60plus'
#
# # Add tripID variable to it
# fnotrips$TripID <- c(max(baseline$TripID) + 1:nrow(fnotrips))
#
# rm(shortwalks, df)
#
# #Sample before running scenarios -
#
# # Removed NAs from the data.frame
# hsematch <- readRDS(system.file("extdata", "hse-nts_match_v2.rds", package = "scenic"))
# #names(hsematch)[c(1,7,8)] <- c('ID', 'health_mmets', 'physical_activity_mmets')
# #hsematch = hsematch[, c(1,7,8)]
#
# #next line commented as people w/o trips already in hse
# #hsematch <- rbind(hsematch, subset(fnotrips, select = c(ID, health_mmets, physical_activity_mmets)))
#
# # Remove health_mmets and physical_activity_mmets from fnotrips
# fnotrips$health_mmets <- NULL
# fnotrips$physical_activity_mmets <- NULL
#
# baseline <- rbind.fill(baseline, fnotrips)
#
# # Read nts age group lookup table
# ag_lookup <- read.csv(system.file("extdata", "nts-adjusted-age-groups.csv", package = "scenic"), header = T, as.is = T)
#
# # Create a new variable 'age_group' for baseline, which converts numeric age categories into age ranges
# baseline$age_group <- ag_lookup$age[match(baseline$Age_B01ID, ag_lookup$nts_group)]
# rm(ag_lookup)
#
#
# baseline[is.na(baseline)] <- 0
# baseline$TripID  <- as.numeric(factor(baseline$TripID))
#
#
# #hsematch <- hsematch[,c(8,9)]  #keep only first and last column > IndivID, mMETs
# hse1 <- setDT(hsematch)[,if(.N<1) .SD else .SD[sample(.N,1,replace=F)],by=ID]
# hse1 = as.data.frame(hse1)
#
# #add times cols. to baseline (used for travel time after cycle switch takes place)
# TripTotalTime1 <- 0
# TripTravelTime1 <- 0
# baseline <- cbind(baseline,TripTotalTime1,TripTravelTime1)
# baseline$TripTotalTime1 <- baseline$TripTotalTime
# baseline$TripTravelTime1 <-baseline$TripTravelTime
#
# #add mmets column to baseline (& save for having total mmets)
# baseline <-inner_join(baseline,hse1,by='ID')
# rm(hse1, hsematch)
#
# #randcycle (used later to calculate if people are cyclists), add col. [prob]
# randcycle <- runif(length(unique(baseline$ID)))
# randcycle <- data.frame(ID=unique(baseline$ID),prob=randcycle)
# baseline <- inner_join(baseline,randcycle,by='ID')
# rm(randcycle)
#
# baseline$prob[baseline$TravDay==0 ] <- 0
#
#
# #keep bl as backup for future scenarios core values
# bl <- baseline
#
# #save PROCESSED baseline in main folder
# #saveRDS(bl,file='bl2014_p_v2.rds')
# rm(bl)
#
# #write.csv(bl,file='bl2012_18_84ag_sw_reduced.csv', row.names=F)
#
#
# ###################################  START CALCULATIONS on BASELINE #############################
#
# #nos. for baseline
# carMiles0 <- sum(baseline[baseline$MainMode_B04ID %in% c(3,4,5,12),'TripDisIncSW'])
# carMiles0 <- round(carMiles0,1)
# METh0 <- round(sum(baseline$METh),1)
# MMETh0 <- round(sum(baseline$MMETh),1)
# # Miles to Kilometres, Grams to metric tonnes, 0.0001
# # Using new Christian's average CO2 value of 0.31 grams
# CO20 <- round(carMiles0 * 1.61 * (3.1 / 1.61) * 1e-1,2)   #(in Kg)
#
# # df <- data.frame()
#
# #i <- c(2,4,8,16,32,64)
#
# # DF which main role is to store info of every case in which Observed > DP. Used in UI to filter out mentioned cases (used in 1_flowgram2-2012)
#
# directProbCasesAboveGivenPerc <- data.frame(MS = numeric(0),
#                                             ebikes = numeric(0),
#                                             equity = numeric(0),
#                                             region = numeric(0))
#
# # directProbs values
#
# directProbs <- c(0.05, 0.1, 0.25, 0.5, 0.75, 1)
#
# # Removing TDR
# # TDR
# # j  <- c(1,0.9,0.8,0.7)
#
# m <- c(0,1)   #ebikes
# n <- c(0,1)   #equity
# num = 1
#
# #keep only used variables
# baseline <-baseline[ , c('Age_B01ID', 'Sex_B01ID', 'HHoldGOR_B02ID', 'CarAccess_B01ID',
#                          'NSSec_B03ID', 'IndIncome2002_B02ID', 'EthGroupTS_B02ID', 'TripID', 'ID',
#                          'MainMode_B03ID', 'MainMode_B04ID', 'MainMode_B11ID',
#                          'TripTotalTime', 'TripTravTime', 'TripDisIncSW',
#                          'agesex', 'Cycled', 'METh', 'MMETh', 'cyclist',
#                          'Age', 'Sex', 'age_group',
#                          'TripTotalTime1', 'TripTravelTime1',
#                          'health_mmets', 'physical_activity_mmets', 'prob')]
#
#
# listOfScenarios <- list()
#
# # for (ebikes in m) {
# #   for (equity in n) {
# #     for  (MS in directProbs) { # all occurences of MS should be replaced
# #       cat(ebikes, equity, MS, "\n")
# #       scenario_name <- paste("MS",MS,"_ebik",ebikes,"_eq" ,equity,sep="")
# #       #assign(scenario_name,flowgram(baseline, MS,ebikes,equity, pcycl_baseline))
# #       tempSc <- flow_gram(baseline, MS,ebikes,equity, pcycl_baseline)
# #       #saveRDS(tempSc, paste0('./temp_data_folder/output/repo_version/', scenario_name, '.rds'))
# #
# #       listOfScenarios[[num]] <- scenario_name
# #       num <- num + 1
# #     }
# #   }
# # }  #j-i-m-n loop
#
# cat('All done! \n')
```