### 2017-02-13 R/cambridge.Rmd

# Integrated Transport and Health Impacts Model

## ITHIM Package
<!-- This document will eventually serve as a compendium to the R package
_ITHIM_.  It will describe how to use the package to implement the
ITHIM model, as well as a detailed description of the model itself.

As a working document it will serve as a place for SGY to post
results, questions, interpreations, notes, etc.-->

```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
library("devtools")
install_github("syounkin/ITHIM", ref="devel")
#install("~/ITHIM/")
#document("~/ITHIM/")
#install("~/ITHIM/")
library("ITHIM")
```

```{r misc, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

### Create the ITHIM object

#### Method 1: Using Input Files

We begin by creating the ITHIM object from which both the baseline and
scenarios are derived.  The necessary input parameters are contained
in 3 files, listed below.

```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}

fileNamesList <- list(
    activeTransportTimeFile = "~/GHI/misc/activeTransportTime.csv",
    roadInjuriesFile = "~/GHI/misc/stata19.csv",
#    roadInjuriesFile = "~/GHI/misc/roadInjuries.csv",
    GBDFile = "~/GHI/misc/gbdEngland_20170209.csv"
#    GBDFile = "~/GHI/misc/gbd.csv"
)

ITHIM <- createITHIM(fileNamesList)
#ITHIM <- createITHIM()
```

The file activeTtransportTime.csv contains means (absolute) for
walking and cycling times (min/week), stratified by age and sex.
Links to sample files can be found below.

http://ithim.ghi.wisc.edu/activeTransportTime.csv
http://ithim.ghi.wisc.edu/roadInjuries.csv
http://ithim.ghi.wisc.edu/gbd.csv

Let's verify that our parameters have been set properly.

```{r misc14, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getWalkTime(ITHIM)
getCycleTime(ITHIM)
getRoadInjuries(ITHIM)
head(getGBD(ITHIM))
```

## Archive

The ITHIM class, which defines an ITHIM object, contains three slots.  Two of the slots are used for the physical activity component and the third is simply a container for all of the parameters in the model.  This container is the ParameterSet class.

```{r roadInjuries11, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getParameterSet(ITHIM)
```

## Road Injury Model
We inspect the road injury counts that are used by default.
```{r misc4, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
RI <- getRoadInjuries(ITHIM)
class(RI)
names(RI)
RI$FatalLocal
```

We do not yet have default values for the distance by road type
matrices.

```{r misc3, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getDistRoadType(ITHIM)
```

So, we have to update manually.

```{r roadInjuries, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
scenario <- readRDS(file = "~/Downloads/distByRoadTypeScenario4.rds")
base <- readRDS(file = "~/Downloads/distByRoadTypeBaseline.rds")
ITHIM.baseline <- update(ITHIM, list(distRoadType = base))
ITHIM.scenario <- update(ITHIM, list(distRoadType = scenario))
getDistRoadType(ITHIM.baseline)
getDistRoadType(ITHIM.scenario)
```

Now that we have the distances by road type we may compute the
scenario injury multipliers and compute the scenario road injury
estimates.

```{r roadInjuries0, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
```

The final ingredient is the GBD set of parameters.  We have a file
stored within the ITHIM package from which we may read in these
values.  We update the ITHIM objects as follows.

```{r roadInjuries9, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
#ITHIM.baseline <- update(ITHIM.baseline, list(GBD = readGBD(file = "~/GHI/misc/gbdEngland.csv")))
#ITHIM.scenario <- update(ITHIM.scenario, list(GBD = readGBD(file = "~/GHI/misc/gbdEngland.csv")))
```

```{r roadInjuries10, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getParameterSet(ITHIM.baseline)
```

```{r roadInjuries3, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
(deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta))
```

## Safety in Numbers

The point of all of this class and method business is to make the code for analyses tidy and human-readable.  Here is an example in which we vary the safety in numbers parameters.

```{r roadInjuries4, message = FALSE, , dev=c('png'), eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}

deltaDeathsVec <- c()

for( sin in sinVec <- seq(0,1,length.out = 10) ){
    ITHIM.baseline <- update(ITHIM.baseline, list(safetyInNumbers = sin*matrix(1,nrow = 8, ncol = 2)))
    ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
    RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
    deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta) # Not human-readable enough, should improve
    deltaDeathsVec <- c(deltaDeathsVec, deltaDeaths)
}

plot(sinVec, deltaDeathsVec, type = "b", pch = 20)

```

## Uncertainty in Road Injury Estimates

```{r roadInjuries6, message = FALSE, , dev=c('png'), eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}

deltaDeaths.vec <- c()
for( i in 1:50 ){

  simRI <- lapply(getRoadInjuries(ITHIM.baseline), function(x){
    lambda <- c(as.matrix(x))
    simCount <- matrix(mapply(function(y){rpois(1,y)},y=lambda), nrow = nrow(x), ncol = ncol(x))
    dimnames(simCount) <- dimnames(x)
    return(as.data.frame(simCount))
  })
  ITHIM.baseline.sim <- update(ITHIM.baseline, list(roadInjuries = simRI))
  ITHIM.scenario.sim <- updateRoadInjuries(ITHIM.baseline.sim, ITHIM.scenario)

  RIburden <- computeRoadInjuryBurden(ITHIM.baseline.sim, ITHIM.scenario.sim)
  deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
  deltaDeaths.vec <- c(deltaDeaths.vec, deltaDeaths)
}
hist(deltaDeaths.vec)
```