### 2017-02-23 R/RoadFatalities.Rmd
Cleaned up R directory

## Road Fatalities

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
#install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
library("ggplot2")
library("devtools")
library("reshape2")
library("plyr")
library("dplyr")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
#document("~/ITHIM/")
#install("~/ITHIM/")
install_github("syounkin/ITHIM", ref="devel")
install_github("smbache/magrittr", ref="master")
#library("magrittr")
library("ITHIM")
fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

```{r roadFat, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
pedDeath <- readRDS(file = "./data/BikePedDeathModelData.rds")
pedDeath <- subset(pedDeath, urbSubRur == "Urban")
pedDeath <- within(pedDeath, pedDeathLogical <- pedDeaths0913 > 0)
pedDeath <- within(pedDeath,{
  deathsTotal <- pedDeaths0913 + bikeDeaths0913
  deathsTotalLogical <- deathsTotal > 0
  pctATToWork <- pctWalkToWork + pctBikeToWork
AT <- ifelse(pctATToWork > 0  & pctATToWork <= 0.10 , "low" ,
ifelse(pctATToWork > 0.10  & pctATToWork <= 0.50 , "medium" ,
ifelse(pctATToWork > 0.50  & pctATToWork <= 1 , "high" , NA)))

})

D <- melt(pedDeath)

chisq.test(with(pedDeath, table(deathsTotalLogical, activeTract)))

p <- ggplot(pedDeath, aes(pedDeaths0913, ..density..,colour = AT))
p #+ geom_freqpoly(binwidth = 10) #+ facet_grid( SEX ~ .)

ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) +

```

AT <- ifelse( ( pctWalkToWork + pctBikeToWork ) > 0  &  ( pctWalkToWork + pctBikeToWork ) <= 0.10 , "low" ,
ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.10  &  ( pctWalkToWork + pctBikeToWork ) <= 0.50 , "medium" ,
ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.50  &  ( pctWalkToWork + pctBikeToWork ) <= 1 , "high" , NA)))