### 2017-02-27 R/RoadFatalities.Rmd
Created US National Road Injury estimates

## Road Fatalities

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
#install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
library("ggplot2")
library("devtools")
library("reshape2")
library("plyr")
library("dplyr")
library("tidyr")
#install_github("smbache/magrittr", ref="master")
#library("magrittr")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
install("~/ITHIM/")
#install_github("syounkin/ITHIM", ref="devel")
library("ITHIM")
fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

Note that I edited the file 'Countyavg_01-05.csv' by hand a little.
The changes were simple, e.g., column names and capitilization.

```{r foobar, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
RIbyCounty <- read.csv(file = "~/GHI/misc/Countyavg_01-05.csv", stringsAsFactors = TRUE)
```

We create a national road injuries estimate by adding up all of the
county averages.  We do that using the R package `tidyr` and its
piping mechanism `%>%`.  This is then written to the default file in
`inst` and read in with `createITHIM` using no arguments (because its
the default file).

```{r foo, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
RINational <- RIbyCounty %>%
  gather(strikingMode, count, -(1:5)) %>%
  complete(severity, roadType, victimMode,strikingMode) %>%
  group_by(severity,roadType,victimMode,strikingMode) %>%
  summarise(value = sum(count, na.rm = TRUE)) %>%
  filter(strikingMode != "nov")

write.csv(RINational, file = "~/ITHIM/inst/roadInjuriesUS.csv", quote = FALSE, row.names = FALSE)
```
```{r foo2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
install("~/ITHIM/")
ITHIM <- createITHIM()

getRoadInjuries(ITHIM)
```
Rows are stiking mode and columns are victim mode.

```{r roadFat, eval = FALSE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, echo = FALSE, warning = FALSE}
pedDeath <- readRDS(file = "./data/BikePedDeathModelData.rds")
pedDeath <- subset(pedDeath, urbSubRur == "Urban")
pedDeath <- within(pedDeath, pedDeathLogical <- pedDeaths0913 > 0)
pedDeath <- within(pedDeath,{
  deathsTotal <- pedDeaths0913 + bikeDeaths0913
  deathsTotalLogical <- deathsTotal > 0
  pctATToWork <- pctWalkToWork + pctBikeToWork
AT <- ifelse(pctATToWork > 0  & pctATToWork <= 0.10 , "low" ,
ifelse(pctATToWork > 0.10  & pctATToWork <= 0.50 , "medium" ,
ifelse(pctATToWork > 0.50  & pctATToWork <= 1 , "high" , NA)))

})

D <- melt(pedDeath)

chisq.test(with(pedDeath, table(deathsTotalLogical, activeTract)))

p <- ggplot(pedDeath, aes(pedDeaths0913, ..density..,colour = AT))
p #+ geom_freqpoly(binwidth = 10) #+ facet_grid( SEX ~ .)

ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) +

#AT <- ifelse( ( pctWalkToWork + pctBikeToWork ) > 0  &  ( pctWalkToWork + pctBikeToWork ) <= 0.10 , "low" ,
#ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.10  &  ( pctWalkToWork + pctBikeToWork ) <= 0.50 , "medium" ,
#ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.50  &  ( pctWalkToWork + pctBikeToWork ) <= 1 , "high" , NA)))

```