### 2017-02-28 R/tutorial.Rmd

# ITHIM Package Tutorial

```{r rm, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
```
To begin using ITHIM one first must install it.  The package devtools
makes this is easy.

## Install and Load ITHIM

```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("devtools")
install_github("syounkin/ITHIM", ref="master")
library("ITHIM")
```

## Creating an ITHIM object

The ITHIM package uses the ITHIM class to facilitate...

### Method 1: Using Input Files

We begin by creating an ITHIM object to represent the baseline model.
Currently ITHIM reads three files for the necessary input parameters.
Sample files are listed below.

```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
fileNamesList <- list(
    activeTransportTimeFile = "~/ITHIM/inst/activeTransportTime.csv",
    roadInjuriesFile = "~/ITHIM/inst/roadInjuries.csv",
    GBDFile = "~/ITHIM/inst/gbd.csv"
)
```

These files can be found in the master branch of ITHIM repository.

* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/master/inst/activeTransportTime.csv)
* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)
* [gbd.csv](https://github.com/syounkin/ITHIM/blob/master/inst/gbd.csv)

Using the list of three file names we can create the ITHIM object using `createITHIM`.

```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline <- createITHIM(fileNamesList)
```

#### Active Transport File Format
```
ageClass,sex,mode,value
ageClass1,M,walk,10
ageClass2,M,walk,9
...
ageClass7,F,cycle,15
ageClass8,F,cycle,13
```
The rows must be ordered by ageClass.
#### Road Injuries File Format
```
severity,roadType,strikingMode,victimMode,value
Fatal,Local,walk,walk,0
Fatal,Local,cycle,walk,13
...
Serious,Highway,LGV,NOV,0
Serious,Highway,mbike,NOV,0
```
ebikes are not included
#### GBD File Format
The GBD file format has not been normalized like the others yet.
```
disease,sex,ageClass,dproj,yll,yld,daly
BreastCancer,M,ageClass1,0,0,0,0
BreastCancer,M,ageClass2,0,0,0,0
...
RTIs,F,ageClass7,1486.647682,21930.32348,7921.96966,29852.3031
RTIs,F,ageClass8,1957.865757,10720.47109,5222.446302,15942.95826
```
RTIs stands for road traffic injuries.

```{r misc22, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
purl("./tutorial.Rmd","./tutorialScript.R", documentation = 0)
```

### Method 2: Using the `update` Method

Suppose we wanted to create a scenario in which the mean cycling time
and mean walking time are doubled.  We need only to change two means
in the model and nothing else.  To do so we could avoid creating input
files by simply updating the baseline object with a few new
parameters.

First we need to know the name of the parameters we want to update.
Currently the list of parameters resides in the help page for the
ITHIM class.  Alternatively one can inspect the
slots directly with `slotNames`.

```{r misc24, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = TRUE}
help("ITHIM")
```
or
```{r misc124, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
slotNames(getParameterSet(ITHIM.baseline))
```

Now we need to retrieve the walking and cycling means from the
baseline model.  To do so we use the method `getMeans`.  To find a
list of available methods use `getMethodsITHIM` with no arguments.

```{r misc23, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getMethodsITHIM()
meanVec <- getMeans(ITHIM.baseline)
meanVec
meanWalkingTime <- meanVec$walk; meanCyclingTime <- meanVec$cycle
scenarioParameters <- list(muwt = 2*meanWalkingTime, muct = 2*meanCyclingTime)
scenarioParameters
```

Now we use the list of parameters and the baseline ITHIM object to
create the scenario ITHIM object.

```{r misc250, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.scenario <- update(ITHIM.baseline, scenarioParameters)
```

Let's inspect it to be sure.

```{r misc351, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getMeans(ITHIM.scenario)
```

## Physical Activity Component

Once the baseline and scenario ITHIM objects are defined, we may
compute the change in burden due to physical activity as follows.

```{r misc25, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "total", bur = "deaths")
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "BreastCancer", bur = "deaths")
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "deaths")
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "daly")
```

## Appendix

```{r misc350, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline
ITHIM.scenario
```
### Session Info
```{r sessionInfo, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
sessionInfo()
```