### 2017-03-16 R/tutorial.Rmd
Update to tutorial

# ITHIM Package Tutorial

```{r rm, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
```
To begin using ITHIM one first must install it.  The package
_devtools_ makes this is easy.  Install _devtools_ using
`install.packages("devtools")` or some other way in _RStudio_.

## Install and Load ITHIM

We use the function `install_github` to install the _master_ branch of
the _ITHIM_ package repository.

```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("devtools")
install_github("syounkin/ITHIM", ref="master")
library("ITHIM")
```

## Creating a Baseline ITHIM object

We begin by creating an ITHIM object to represent the baseline model.
Currently _ITHIM_ can read up to three files containing input
parameters, but it is not necessary to provide all three.  Two sample
files are listed below that contain model parameters for the active
transport component of _ITHIM_.

```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
activeTransportFile <- "~/ITHIM/inst/activeTransportTime.csv"
GBDFile <- "~/ITHIM/inst/gbd.csv"
```

For a description of the file format see the help page for
`createITHIM` using `help("createITHIM")`.

```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile,
                              GBDFile = GBDFile)
```

* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/master/inst/activeTransportTime.csv)
* [gbd.csv](https://github.com/syounkin/ITHIM/blob/master/inst/gbd.csv)

### File Formats
#### Active Transport File Format

The active transport file contains mean walking and cycling times
stratified by sex and _ITHIM_ age class.  The default file is located
in _inst/_.

```
ageClass,sex,mode,value
ageClass1,M,walk,21.47
ageClass2,M,walk,53.11
ageClass3,M,walk,50.13
...
ageClass6,F,cycle,1.53
ageClass7,F,cycle,0.53
ageClass8,F,cycle,0.23
```

Notes about file format:
* The rows must be ordered by ageClass.
* The mean values are in minutes per week

#### Disease Burden File Format

The disease burden file contains estimates for disease burden
stratified by _ITHIM_ age class and sex.

```
disease,sex,ageClass,burdenType,value
BreastCancer,M,ageClass1,dproj,0
BreastCancer,M,ageClass2,dproj,0
BreastCancer,M,ageClass3,dproj,0
...
RTIs,F,ageClass6,daly,49108.81151
RTIs,F,ageClass7,daly,29852.3031
RTIs,F,ageClass8,daly,15942.95826
```

Notes about file format:
* RTIs stands for road traffic injuries
* Only _ITHIM_ diseases can be included, namely "BreastCancer",
  "ColonCancer", "CVD", "Dementia", "Diabetes", "Depression" and "RTIs"

```{r misc22, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
purl("./tutorial.Rmd","./tutorialScript.R", documentation = 0)
```

#### Road Injuries File Format

If road injury data is present it can be included in the same manner.

```{r misc123, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
roadInjuriesFile <- "~/ITHIM/inst/roadInjuries.csv"
```

```{r misc220, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile,
                              GBDFile = GBDFile,
                              roadInjuriesFile = roadInjuriesFile)
```

* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)

```
severity,roadType,strikingMode,victimMode,value
Fatal,Local,walk,walk,0
Fatal,Local,cycle,walk,13
Fatal,Local,bus,walk,213
...
Serious,Highway,HGV,NOV,0
Serious,Highway,LGV,NOV,0
Serious,Highway,mbike,NOV,0
```

Notes about file format:
* ebikes are not included

## Creating Scenarios Using the `update` Method

Suppose we wanted to create a scenario in which the mean cycling time
and mean walking time are doubled.  We need only to change two means
in the model and nothing else.  To do so we could avoid creating input
files by simply updating the baseline object with a few new
parameters.

First we need to know the name of the parameters we want to update.
Currently the list of parameters resides in the help page for the
ITHIM class (`help("ITHIM")`).  Alternatively one can use
`getParameterNames`.

```{r misc124, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getParameterNames(ITHIM.baseline)
```

Now we need to retrieve the walking and cycling means from the
baseline model.  To do so we use the method `getMeans`.  To find a
list of available methods use `getMethodsITHIM` with no arguments.

```{r misc23, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getMethodsITHIM()
meanVec <- getMeans(ITHIM.baseline)
meanVec
meanWalkingTime <- meanVec$walk; meanCyclingTime <- meanVec$cycle
scenarioParameters <- list(muwt = 2*meanWalkingTime, muct = 2*meanCyclingTime)
scenarioParameters
```

Now we use the list of parameters and the baseline ITHIM object to
create the scenario ITHIM object.

```{r misc250, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.scenario <- update(ITHIM.baseline, scenarioParameters)
```

Let's inspect it to be sure.

```{r misc351, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getMeans(ITHIM.scenario)
```

## Physical Activity Component

Once the baseline and scenario ITHIM objects are defined, we may
compute the change in burden due to physical activity as follows.

```{r misc25, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "all", bur = "deaths")
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "BreastCancer", bur = "deaths")
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "deaths")
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "daly")
```

## Built-In Example

An example is contained within the _ITHIM_ package.

```{r misc225, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
example("ITHIM")
```

## Appendix

### Session Info & Package Test
```{r sessionInfo, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("testthat")
test_package("ITHIM")
sessionInfo()
```