### 2017-06-02 R/oregon.Rmd
Updates to oregon.Rmd

# ITHIMr Demonstration for Portland, Oregon

```{r rm, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
```

## Install and Load ITHIM

The package _devtools_ makes installation of the _ITHIM_ package easy.
Install _devtools_ using `install.packages("devtools")` or some other
way in _RStudio_.  We use the function `install_github` from
_devtools_ to install the _devel_ branch of the _ITHIM_ package
repository.

```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("devtools")
install_github("ITHIM/ITHIM", ref="devel")
```

```{r libs2, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("ITHIM")
library("tidyverse")
```

```{r figwidth, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
fig.width <- 24 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

## Create a Baseline ITHIM object

We begin by creating an ITHIM object to represent the baseline model.
Currently _ITHIM_ can read up to three files containing input
parameters, but it is not necessary to provide all three.  Two sample
files are listed below that contain model parameters for the active
transport component of _ITHIM_.

Sample Files
* [activeTransportTime.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/activeTransportTime.csv)
* [gbd.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/gbd.csv)

The file_activeTransprtTime.csv_ contains weekly mean travel times in
minutes, stratified by age, sex and mode (walk,cycle). Currently eight
age classes are used for the _ITHIM_ model.  For a more detailed
description of the file format see the help page for `createITHIM`
using `help("createITHIM")`.

Please update the path and names in the code below to match the csv files you wish to use.

```{r misc320, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline <- createITHIM(activeTransportFile = "~/Downloads/activeTransportTime.csv",
                              GBDFile = "~/Downloads/gbd.csv")
```

The contents of the _ITHIM_ object are displayed in summary when the
object is printed.  See below.

```{r show, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline
```

Currently the list of internal model parameters resides in the help page for the
ITHIM class (`help("ITHIM")`).  Alternatively one can use the method
`getParameterNames` to display the names.

```{r misc124, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getParameterNames(ITHIM.baseline)
```

If we wanted to examine the overal means in the model we could use the
method `getMeans`, which returns a vector of means.  To find a list of
available methods use `getMethodsITHIM` with no arguments.

```{r misc23, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getMethodsITHIM()
```

We'll store the vector of means in a variable called _meanVec_ and then view that vector.

```{r misc322, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
meanVec <- getMeans(ITHIM.baseline)
meanVec
```

## Creating a Scenario ITHIM object

Suppose we wanted to create a scenario in which the mean cycling time
and mean walking time are doubled.  We simply need to replace the
active transport file with one that has the desired values.

Sample Files
* [activeTransportTime_scenario.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/activeTransportTime_scenario.csv)

Please update the path and names in the code below to match the csv
files you wish to use.

```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.scenario <- createITHIM(activeTransportFile = "~/Downloads/activeTransportTime_scenario.csv",
                              GBDFile = "~/Downloads/gbd.csv")
```

```{r misc7620, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
modelMeans <- rbind(cbind(rbind(getWalkTime(ITHIM.baseline), getCycleTime(ITHIM.baseline), getNonTravelMETs(ITHIM.baseline)),vision = "baseline"), cbind(rbind(getWalkTime(ITHIM.scenario), getCycleTime(ITHIM.scenario), getNonTravelMETs(ITHIM.scenario)),vision = "scenario"))

p <- ggplot(modelMeans, aes(x = ageClass, y = mean, fill = vision))
p + geom_bar(stat = "identity", position = "dodge") + facet_grid(type ~ sex, scales = "free_y")  + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Physical Activity Component

Once the baseline and scenario ITHIM objects are defined, we may
compute the change in burden due to physical activity as follows. We
provide a few examples to give you an idea of options for diseases and
burden.

```{r misc25, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "all", bur = "deaths")
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "BreastCancer", bur = "deaths")
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "deaths")
deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "daly")
```

## Creating Scenarios Using the `update` Method

In the above example we needed only to change two means in the model
and nothing else.  To do so we could avoid creating input files by
simply updating the baseline object with a few new parameters.  (note:
this is the overall mean for the whole population. From this, _ITHIM_
will create means for each age & gender strata)

We can use the mode-specific means from the _meanVec_ vector and
assign them to their own variables.

```{r misc3322, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
meanWalkingTime <- meanVec$walk
meanCyclingTime <- meanVec$cycle
```

Next we'll create a list of the parameters we wish to change in our
scenario. Recall, we're interested in a scenario that has double the
mean walking time and double the mean cycling time of the baseline
scenario.

```{r misc3242, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
scenarioParameters <- list(muwt = 2*meanWalkingTime, muct = 2*meanCyclingTime)
scenarioParameters
```

_ITHIM_'s comparative risk assessment framework uses two models to describe the potential health impacts for small differences between them. In our case the two models differ in the active travel times of the population.  Above we created the baseline model, and now we use the list of changed parameters to `update` the baseline ITHIM object. This method creates a copy of the baseline ITHIM object with the changes we'd like to see in the new ITHIM object. We call this new object the scenario.

```{r misc250, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.scenario <- update(ITHIM.baseline, scenarioParameters)
```

Let's inspect the scenario ITHIM object to be sure it includes the changes we expect.

```{r misc351, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getMeans(ITHIM.scenario)
```

## Creating an ITHIM object for Portland (MSA 6442)

### Active Transport Means
We use the National Household Transportation Survey (NHTS) data stored
as a _TravelSurvey_ R object to estimate mean walk and cycle times.

```{r oregon23, eval = TRUE, echo = FALSE, message = FALSE, warnings = FALSE, results = "hide", fig.width = fig.width, fig.asp = 1/phi}
install_github("syounkin/TravelSurvey", ref="master")
library("TravelSurvey")
NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
```

```{r oregon333, eval = TRUE, echo = FALSE, fig.width = fig.width, fig.asp = 1/phi}
NHTS.df <- within(NHTS.df, {
    id <- paste0(houseID,"-",subjectID)
    duration <- ifelse(is.na(duration), 0, duration)
    mode <- factor(mode, levels = c("walk","cycle","other"))
    age <- factor(age, levels = c("00-04","05-14","15-29","30-44","45-59","60-69","70-79","80+"))
    sex <- factor(sex, levels = c("M","F"))
})
```

```{r oregon2, eval = TRUE, echo = TRUE, fig.width = fig.width, fig.asp = 1/phi}
sampleSize <- NHTS.df %>% filter(location == "6442") %>% group_by(sex,age) %>% summarise(n=n_distinct(id))

tripDuration <- NHTS.df %>% filter(location == "6442") %>% group_by(mode,sex,age) %>% summarise(total=sum(duration))

activeTravelOregon <- right_join(sampleSize,tripDuration, by = c("sex","age")) %>% ungroup() %>% mutate(mean = total/n*7) %>% select(mode,sex,age,mean) %>% complete(mode,sex,age) %>% filter(mode %in% c("walk","cycle")) %>% as.data.frame()

activeTravelOregon
```

### Disease Burden

```{r gbd, eval = TRUE, echo = TRUE, fig.width = fig.width, fig.asp = 1/phi}
```

## Appendix

### Built-In Example

An example is contained within the _ITHIM_ package.

```{r misc225, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
example("ITHIM")
```

### File Formats
#### Active Transport File Format

The active transport file contains mean walking and cycling times
stratified by sex and _ITHIM_ age class.  The default file is located
in _inst/_.

```
ageClass,sex,mode,value
ageClass1,M,walk,21.47
ageClass2,M,walk,53.11
ageClass3,M,walk,50.13
...
ageClass6,F,cycle,1.53
ageClass7,F,cycle,0.53
ageClass8,F,cycle,0.23
```

Notes about file format:
* The rows must be ordered by ageClass.
* The mean values are in minutes per week

#### Disease Burden File Format

The disease burden file contains estimates for disease burden
stratified by _ITHIM_ age class and sex.

```
disease,sex,ageClass,burdenType,value
BreastCancer,M,ageClass1,dproj,0
BreastCancer,M,ageClass2,dproj,0
BreastCancer,M,ageClass3,dproj,0
...
RTIs,F,ageClass6,daly,49108.81151
RTIs,F,ageClass7,daly,29852.3031
RTIs,F,ageClass8,daly,15942.95826
```

Notes about file format:
* RTIs stands for road traffic injuries
* Only _ITHIM_ diseases can be included, namely "BreastCancer",
  "ColonCancer", "CVD", "Dementia", "Diabetes", "Depression" and "RTIs"

#### Road Injuries File Format

* [roadInjuries.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/roadInjuries.csv)

```
severity,roadType,strikingMode,victimMode,value
Fatal,Local,walk,walk,0
Fatal,Local,cycle,walk,13
Fatal,Local,bus,walk,213
...
Serious,Highway,HGV,NOV,0
Serious,Highway,LGV,NOV,0
Serious,Highway,mbike,NOV,0
```

Notes about file format:
* ebikes are not included

### Session Info & Package Test
```{r sessionInfo, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("testthat")
test_package("ITHIM")
sessionInfo()
```