### 2017-06-07 R/msa.Rmd

foo <- read.csv(file="~/GHI/data/join2.csv", stringsAsFactors = FALSE)

msaTest <- 6442

msaMatrix <- foo %>% filter(msaNHTS == msaTest) %>%
select(stateFIPS,countyFIPS)

getPopulation(state=msaMatrix$stateFIPS, county =
msaMatrix$countyFIPS)