### 2017-06-08 R/msa.Rmd
Created msa.Rmd and added msa to Makefile

# MSA Population Size (F)

```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
library("devtools")
install_github("ITHIM/ITHIM", ref="devel")
library("ITHIM")
```
## List of Population Size Matrices

```{r msa, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
msaIDFile <- read.csv(system.file("join2.csv", package = "ITHIM"), stringsAsFactors = FALSE)
msaIDs <- msaIDFile$msaNHTS %>% unique
msaIDs <- msaIDs[!is.na(msaIDs)] %>% as.list
names(msaIDs) <- paste0("MSA-NHTS-", msaIDs)
populationList <- getPopulation(msaIDs)

length(populationList)
names(populationList)
head(populationList, 3)
```

## Top Three MSAa

Let's inspect the top three MSAs by population size.

```{r msa2, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
totalPop <- unlist(lapply(populationList,sum))
head(totalPop[order(totalPop, decreasing = TRUE)],3)
msaIDFile %>% filter(msaNHTS %in% c(4472,5602,1602)) %>% select(msaName,msaNHTS) %>% unique
```

Notice the non-uniqueness of `msaNHTS` in the data.