### 2017-06-09 R/msa.Rmd
Update to TravelSurvey.Rmd and msa.Rmd

# MSA Population Size (F)

```{r msalibs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
library("devtools")
install_github("ITHIM/ITHIM", ref="devel")
library("ITHIM")
```
## List of Population Size Matrices

```{r msamsa, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
msaIDFile <- read.csv(system.file("join2.csv", package = "ITHIM"), stringsAsFactors = FALSE)
msaIDs <- msaIDFile$msaNHTS %>% unique
msaIDs <- msaIDs[!is.na(msaIDs)] %>% as.list
names(msaIDs) <- paste0("MSA-NHTS-", msaIDs)
populationList <- getPopulation(msaIDs)

length(populationList)
names(populationList)
head(populationList, 3)

saveRDS(populationList, file = "~/GHI/R/data/populationList.rds")
```

## Top Three MSAs

Let's inspect the top three MSAs by population size.

```{r msa2, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
totalPop <- lapply(populationList,sum) %>% unlist
totalPop[totalPop %>% order(., decreasing = TRUE)] %>% head(., 3)
msaIDFile %>% filter(msaNHTS %in% c(4472,5602,1602)) %>% select(msaName,msaNHTS) %>% unique
```

It's L.A. New York and Chicago.  Seems right.

Notice the non-uniqueness of `msaNHTS` in the data.

## MSA ID check

```{r msa3, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
msaIDs <- msaIDFile$msaNHTS %>% unique
msaFromFile <- data.frame(msa = unlist(msaIDs)) %>% filter(!is.na(msa) & !(msa == -1)) %>% unlist

NHTS.df <- readRDS(file = "~/GHI/R/data/NHTS.df.rds")
msaIDs <- NHTS.df$location %>% unique
msaFromNHTS <- data.frame(msa = unlist(msaIDs)) %>% filter(!is.na(msa) & !(msa == -1)) %>% unlist

data.frame(msaFromFile[order(msaFromFile)],msaFromNHTS[order(msaFromNHTS)])
```

Sweet.