### 2017-06-28 R/msa.Rmd

# MSA Population Size (F)

```{r msalibs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
library("devtools")
#install_github("ITHIM/ITHIM", ref="devel")
install("~/ITHIM/")
library("ITHIM")
```
## List of Population Size Matrices

```{r msamsa, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
msaIDFile <- read.csv(system.file("join2.csv", package = "ITHIM"), stringsAsFactors = FALSE)
msaIDs <- msaIDFile$msaNHTS %>% unique
msaIDs <- msaIDs[!is.na(msaIDs)] %>% as.list
names(msaIDs) <- msaIDs
FList <- getPopulation(msaIDs)

length(FList)
names(FList)
head(FList, 3)

saveRDS(FList, file = "~/GHI/R/data/FList.rds")
```

```{r msamsa2, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
NHTS.df <- readRDS(file = "~/GHI/R/data/NHTS.df.rds")
FList <- readRDS(file = "~/GHI/R/data/FList.rds")

formatMeanMatrix <- function(meanTable){
    meanMatrix <- meanTable %>% select(age,sex,mean) %>% ungroup() %>% spread(sex,mean) %>% select(M,F) %>% as.matrix()
    row.names(meanMatrix) <- paste0("ageClass",1:8)
    return(meanMatrix)
}

msaIDs <- unique(NHTS.df$location)
msaIDs <- as.character(msaIDs[msaIDs!="-1"])

sampleSize <- NHTS.df %>% group_by(location,sex,age) %>% summarise(n=n_distinct(id))
tripDuration <- NHTS.df %>% group_by(location,mode,sex,age) %>% summarise(total=sum(duration))
foo <- right_join(sampleSize,tripDuration, by = c("location","sex","age")) %>% ungroup() %>% mutate(mean = total/n*7) %>% select(location,mode,sex,age,mean) %>% complete(location,mode,sex,age)

meanByLocationByMode <- lapply(split(foo,foo$mode), function(x) split(x, x$location))

MwtList <- lapply(meanByLocationByMode$walk, formatMeanMatrix)
MctList <- lapply(meanByLocationByMode$cycle, formatMeanMatrix)
RwtList <- lapply(MwtList, function(x) x/x[3,2])
RctList <- lapply(MctList, function(x) x/x[3,2])

meanWalkList <- mapply(function(x,y) sum(x*y/sum(y), na.rm = TRUE), MwtList[msaIDs], FList[msaIDs], SIMPLIFY = FALSE)
meanCycleList <- mapply(function(x,y) sum(x*y/sum(y), na.rm = TRUE), MctList[msaIDs], FList[msaIDs], SIMPLIFY = FALSE)

cbind(data.frame(walk = unlist(meanWalkList)), data.frame(cycle = unlist(meanCycleList)))

msaID.portland <- "6442"
FList[[msaID.portland]] %>% gather(sex,value,1:2) %>% mutate(ageClass=rep(rownames(FList[[msaID.portland]]),2)) %>% select(ageClass,sex,value) %>% write.csv(., file = "~/ITHIM/inst/F.portland.csv", quote = FALSE, row.names = FALSE)

rbind(MwtList[[msaID.portland]] %>% as.data.frame() %>% gather(sex,value,1:2) %>% mutate(ageClass=rep(rownames(FList[[msaID.portland]]),2)) %>% select(ageClass,sex,value) %>% mutate(mode = "walk"), MctList[[msaID.portland]] %>% as.data.frame() %>% gather(sex,value,1:2) %>% mutate(ageClass=rep(rownames(FList[[msaID.portland]]),2)) %>% select(ageClass,sex,value) %>% mutate(mode = "cycle")) %>% select(ageClass,sex,mode,value) %>% write.csv(., file = "~/ITHIM/inst/activeTransport.portland.csv", quote = FALSE, row.names = FALSE)

```

## Top Three MSAs

Let's inspect the top three MSAs by population size.

```{r msa2, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
totalPop <- lapply(FList,sum) %>% unlist
totalPop[totalPop %>% order(., decreasing = TRUE)] %>% head(., 3)
msaIDFile %>% filter(msaNHTS %in% c(4472,5602,1602)) %>% select(msaName,msaNHTS) %>% unique
```

It's L.A. New York and Chicago.  Seems right.

Notice the non-uniqueness of `msaNHTS` in the data.